Off the beaten path: \\ Gaia reveals GD-1 stars perturbed off of the main stream track

### 2018-04-30 paper/gd1.tex
begin discussion

- novel way to identify members -- helpful that retrograde, but hopefully also applicable to other streams as kinematic important
- gd-1 even longer -- useful for potential constraints --> future

- progenitor? if confirmed, can create more realistic models (so far only modeled the whole stream as just a leading or trailing arm)
- to confirm -- measure actions, which change direction (some bovy paper)?

- no longer simple stream, discuss possible causes
- evaporation?

@@

>>>>

In this work we mapped a cold stellar stream in individual tracers for the first time by combining precise \pans\ photometry with the revolutionary \gaia\ data set.
\gaia\ proper motions were especially helpful for selecting members of the retrograde GD-1 stream, and this success signals a novel way to find members of other structures in the Galactic halo.
In the rest of this section we discuss how the first view of GD-1 in the \gaia\ era is already transforming what we expect to learn about the Galaxy from streams in general, and GD-1 in particular.

The absence of its progenitor has been a long-standing hurdle in dynamical modeling of GD-1, mostly due to the unknown true extent of the stream.
Confirming the progenitor location suggested here, $(\phi_1, \phi_2) = (-13.5^\circ,-0.5^\circ)$, for example by measuring the (canonical) angle variations along the stream \citep{Bovy:2014}, would motivate further searches for stream members along the trailing arm (positive $\phi_1$) to symmetrically match the extent of the leading arm.
The stream length is directly proportional to the information it provides on the underlying gravitational potential \citep{Bonaca:2018}, making GD-1 (even at its current length of $\sim90^\circ$) a top modeling priority to map the inner 30\,kpc of the Galaxy.
Properly accounting for the progenitor would also unleash the full constraining power of detailed density structure within the tails, as advocated by \citet{Kupper:2015}.

@@

* no longer simple stream, discuss possible causes
